Fixing Vulnerabilities Hinders Maintainability

### 2019-02-04 paper/ICPC19/icpc19.tex
issues fixed until S3 and the end of subsection B

Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between two parties. DTLS provides a mechanism for fragmenting a
handshake message over a number of packets, each of which can be transmitted
separately. In the example (Listing~\ref{lst:vuln}), all functions receive as
input the SSL session, \texttt{s}. In the \texttt{dtls1\_reassemble\_fragment}
function the size of the fragment \texttt{msg\_hdr-$>$frag\_off+frag\_len} is
checked against the handshake message size \texttt{msg\_hdr->msg\_len} (line
$13$) and then against the maximum size allowed in a DTLS handshake message for \texttt{s} (line $16$-$21$)\Rui{DTLS1 for s?}. However, it is not verified if \texttt{msg\_hdr->msg\_len} is lower than the size allowed by DTLS. \Rui{$<$- does not parse} Thus, in line
$37$, when \texttt{msg\_len} is allocated to the fragment buffer, more memory
than the allowed might be allocated. According to the commit message, $10$
handshake messages are allowed per DTLS connection. Hence, a remote attacker
could leverage this issue to trigger memory allocations corresponding to large
length values and lead the system to memory exhaustion that could end in a DoS
attack. The code changes performed to fix the vulnerability were the following:

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was
inserted to return the maximum number of bytes allowed in a DTLS
handshake message for \texttt{s}. The minimum is $16$KB but it might be greater if the maximum certificate list size requires it. \Rui{$<$- does not parse} More
specifically, it checks if the maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise, it returns the maximum length. This function
replaces lines $18$-$21$.

@@

\begin{itemize}
    \item \textbf{Using Components with Known Vulnerabilities (21 cases).} The majority of
    software produced today integrates several components, such as libraries and
    frameworks, which may affect the software if developers use vulnerable
    versions - usually known and disclosed somewhere on the Internet.
    \textit{Based on the OWASP Top $10$ of $2017$.}

@@

\emph{Units} are the smallest groups of code that can be maintained and executed independently~\cite{Visser:2016:OREILLY}.

McCabe~\cite{1702388} is a software
metric that calculates the number of linearly independent paths of the source
code. BCH uses this metric to measure the number of branch points in the
\emph{Write Simple Units of Code} guideline. The bar at the bottom of the figure
represents the top $30$ units that violate the guideline, sorted by severity,
The violation severities are indicated using colors, and there is a legend to
help to interpret the colors. The green bar represents the number of compliant
branch points per unit (\emph{at most $5$}), i.e., the units are compliant with
ISO 25010~\cite{iso:2011}. Yellow, orange and red bars represent units that do
not comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint the
compliance thresholds for each severity level. If the green mark is somewhere
in the green bar it means it is compliant with a low level of severity.

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
%
We found that $40.2\%$ of the security refactoring have a negative impact on the
maintainability of open-source software (Figure~\ref{fig:secvsreg}). This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and
	user-friendly~\cite{5287006, 7676144}. Unfortunately, there are several
	vulnerabilities that stay undisclosed. Thus, developers should not trust them
	blindly and check defaults. \Rui{I don't understand the point.}

	\item \textbf{Documentation} of software should provide developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use.

	\item \textbf{Proper Authorization Mechanisms} should be designed and
	implemented. This is one of the most complex security features to implement.
	Some mechanisms might assist authorization maintenance in long run, such as
	the combination of \texttt{HTTP} filters with \texttt{Spring Security}
	annotations. \Rui{not crystal clear -- do you mean programming languages
  should offer these mechanisms? What is that combination that you're talking
  about?}

	\item \textbf{Deprecated code} should be avoided by developers. They should
	integrate tools such as BCH on their projects CI/CD pipelines to help them be
	more aware of the maintainability of their code and issues (e.g., dead and
	duplicated code). \Rui{not clear. deprecated code vs. BCH??}

	\item\textbf{Programming Languages} should provide new mechanisms to easily
	implement security patterns without endangering maintainability. \Rui{merge with
  3rd bullet?}
\end{itemize}

\Rui{Intense, buzz --- this is not a blog. Don't understand the point very
well too.}
We would expect that the negative impact of programming languages on
maintainability would be more intense due to the buzz of the last years
surrounding the idea that poor design of programming languages and lack of using
the best practices lead to more buggy and vulnerable
code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. However,
Figure~\ref{fig:lang_main} shows that only \emph{C} has a negative impact of
more than $50\%$ on software maintainability. We suspect that these values are
the result of projects contributions policies. Nine out of ten projects --- the
equivalent to $31.8\%$ of the commits --- of the top-$10$ of dataset projects
with more contributors have restrict contribution policies regarding code
standards. Some examples are the \texttt{laravel/framework},
\texttt{php/php-src}, \texttt{openssl/openssl} and \texttt{cakephp/cakephp}. For
instance, \texttt{laravel/framework} requires the \texttt{PSR-2
Standard}\footnote{PSR-2 Standard available at
https://www.php-fig.org/psr/psr-2/ (Accessed on \today{})}.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting}
and \emph{Using Components with Known Vulnerabilities} refactorings endure more
cases without any impact in the maintainability. Typically, these types of
vulnerabilities do not need extra lines to be fixed, as shown in
Listing~\ref{lst:fix}. Whereas, \emph{Broken Authentication} and vulnerabilities
such as the one presented in Section~\ref{sec:motivation} responsible by DoS
attacks are more difficult to fix (several code lines deleted and added).
Another example is the Apple authentication flaw discovered in $2017$
(CVE-$2017$-$13872$)\footnote{CVE-$2017$-$13872$ details available at
https://support.apple.com/en-us/HT208315 (Accessed on \today{})}, where any user
could log in as root with an empty password. Patrick Wardle examined the cause
of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at
https://objective-see.com/blog/blog\_0x24.html (Accessed on \today{})} and
conclude that the flaw was due to the introduction of high cyclomatic complexity
in a method used to verify the password. In sum, \emph{Cross-Site Scripting} and
\emph{Using Components with Known Vulnerabilities} should take part of the
developers' preoccupations along with the other patterns that explicitly hinder
maintainability because despite their low complexity these are patterns that
still produce a considerable negative impact on the maintainability. Real
efforts have been made in creating and incorporating coding standards and best
practices in software production. However, security is still far from perfection
and fixing vulnerabilities might hinder software maintainability. Thus, it is
important to provide tool support to developers to help them apply these
patterns without harming software maintainability. \Rui{and redesign programming
languages?}

@@

All the security commits are described in English. This approach discards
security refactorings provided in other languages.\Rui{ehmm, isn't this
related to the first paragraph in the internal threat?}

@@

Many studies have investigated the relationship between refactorings and
software quality. Previous work focused on coupling metrics \Rui{?} has evaluated the
impact of refactorings and exhibited proof that quantifying effectively the
impact of refactoring on maintainability may help to choose the appropriate
refactoring type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of refactorings. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after refactorings. The differences
of maintainability between groups of refactored elements and non-refactored were
analyzed. In addition, the authors conclude that the source code elements
subjected to refactorings had significantly lower maintainability than the
elements not affected. Studying the evolution of maintainability issues during
the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes and faults. The present
work uses a model proposed by BCH and focuses solely on evaluating the impact of
patching vulnerabilities on software maintainability.

@@

This study also shows evidence that there are security patterns that need more attention than
others. \emph{Broken Autentication} issues, \emph{Denial-of-Service} attacks and
\emph{Memory Leaks} endure maintainability in $55.3\%$, $51.3\%$, and $47.5\%$,
respectively. Following the implications of our work, we draw attention for the
need of trustful security frameworks \Rui{?}, code best practices documentation to
produce safe and maintainable code, proper authorization mechanisms design and
implementation, better tools to introduce in the CI/CD pipeline to help software
maintenance and new or improved programming languages to allow developers to
easily implement safety in their software without endangering maintainability.

As future work, this study may be extended in many directions: analyze which
guidelines are more likely to be affected by security-related refactorings;
evaluate the impact of each refactoring pattern on software maintainability
\Rui{?}\Sof{}; investigate if there is any correlation between software
maintainability and popularity of the project; expand our methodology with other
software quality properties; validate these findings with closed or private
software; and, expand this analysis to other quality standards.

>>>>

Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between two parties. DTLS provides a mechanism for fragmenting a
handshake message over a number of packets, each of which can be transmitted
separately. In the example (Listing~\ref{lst:vuln}), all functions receive as
input the SSL session, \texttt{s}. In the \texttt{dtls1\_reassemble\_fragment}
function the size of the fragment \texttt{msg\_hdr-$>$frag\_off+frag\_len} is
checked against the handshake message size \texttt{msg\_hdr->msg\_len} (line
$13$) and then against the maximum size allowed in a DTLS handshake message for \texttt{s} (line $16$-$21$)\Rui{DTLS1 for s?}\Sof{Better?}. However, it is not verified if \texttt{msg\_hdr->msg\_len} is lower than the size allowed by DTLS. \Rui{$<$- does not parse}\Sof{Better?} Thus, in line
$37$, when \texttt{msg\_len} is allocated to the fragment buffer, more memory
than the allowed might be allocated. According to the commit message, $10$
handshake messages are allowed per DTLS connection. Hence, a remote attacker
could leverage this issue to trigger memory allocations corresponding to large
length values and lead the system to memory exhaustion that could end in a DoS
attack. The code changes performed to fix the vulnerability were the following:

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was
inserted to return the maximum number of bytes allowed in a DTLS
handshake message for \texttt{s}. The minimum is $16$KB but it might be greater if the maximum certificate list size requires it. \Rui{$<$- does not parse}\Sof{Better?} More
specifically, it checks if the maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise, it returns the maximum length. This function
replaces lines $18$-$21$.

@@

\begin{itemize}
    \item \Sof{Cases added, check it out}\textbf{Using Components with Known Vulnerabilities (21 cases).} The majority of
    software produced today integrates several components, such as libraries and
    frameworks, which may affect the software if developers use vulnerable
    versions - usually known and disclosed somewhere on the Internet.
    \textit{Based on the OWASP Top $10$ of $2017$.}

@@

One of the guidelines with which the project does not comply is the one presented in the report (Figure~\ref{fig:bchrep}): \emph{Write Simple Units of Code}. SIG defines \emph{Units} as the smallest groups of code that can be maintained and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and constructors in Java).

McCabe~\cite{1702388} is a software
metric that calculates the number of linearly independent paths of the source
code. BCH uses this metric to measure the number of branch points in the
\emph{Write Simple Units of Code} guideline.

\emph{Units} are the smallest sets of code that can be maintained and executed independently~\cite{Visser:2016:OREILLY}. For instance, in Java
units are methods and contructors.

The bar at the bottom of the figure
represents the top $30$ units that violate the guideline, sorted by severity,
The violation severities are indicated using colors, and there is a legend to
help to interpret the colors. The green bar represents the number of compliant
branch points per unit (\emph{at most $5$}), i.e., the units are compliant with
ISO 25010~\cite{iso:2011}. Yellow, orange and red bars represent units that do
not comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint the
compliance thresholds for each severity level. If the green mark is somewhere
in the green bar it means it is compliant with a low level of severity.

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
%
We found that $40.2\%$ of the security refactoring have a negative impact on the
maintainability of open-source software (Figure~\ref{fig:secvsreg}). This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	% \item \textbf{Security Frameworks} need to be secure and
	% user-friendly~\cite{5287006, 7676144}. Unfortunately, there are several
	% vulnerabilities that stay undisclosed. Thus, developers should not trust them
	% blindly and check defaults. \Rui{I don't understand the point.}

	\item \textbf{Documentation} of software should provide developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use.
	\item\textbf{Programming Languages} should provide new mechanisms to easily
	implement security patterns without endangering maintainability. \Rui{merge with
  3rd bullet?}
	Proper Authorization Mechanisms should be designed and
	implemented. This is one of the most complex security features to implement.
	Some mechanisms might assist authorization maintenance in long run, such as
	the combination of \texttt{HTTP} filters with \texttt{Spring Security}
	annotations. \Rui{not crystal clear -- do you mean programming languages
  should offer these mechanisms? What is that combination that you're talking
  about?}

	\item \textbf{Software Developers} should be avoided by developers. They should
	integrate tools such as BCH on their projects CI/CD pipelines to help them be
	more aware of the maintainability of their code and issues.

\end{itemize}

\Rui{Intense, buzz --- this is not a blog. Don't understand the point very
well too.}

% Ray et al. (2017)~\cite{Ray:2017:LSP:3144574.3126905} and
% Berger et al. (2019)~\cite{2019arXiv190110220B}
%
% It would be expected that programming languages would have a more significant impact on the software maintainability according with previous work.
%
% According with the implications of previous work~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}, we would expect that the negative impact of programming languages on
% maintainability would be more significant
% surrounding the idea that poor design of programming languages and lack of using
% the best practices lead to more buggy and vulnerable
% code.

However,
Figure~\ref{fig:lang_main} shows that only \emph{C} has a negative impact of
more than $50\%$ on software maintainability. We suspect that these values are
the result of projects contributions policies. Nine out of ten projects --- the
equivalent to $31.8\%$ of the commits --- of the top-$10$ of dataset projects
with more contributors have restrict contribution policies regarding code
standards. Some examples are the \texttt{laravel/framework},
\texttt{php/php-src}, \texttt{openssl/openssl} and \texttt{cakephp/cakephp}. For
instance, \texttt{laravel/framework} requires the \texttt{PSR-2
Standard}\footnote{PSR-2 Standard available at
https://www.php-fig.org/psr/psr-2/ (Accessed on \today{})}.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site Scripting}
and \emph{Using Components with Known Vulnerabilities} refactorings endure more
cases without any impact in the maintainability. Typically, these types of
vulnerabilities do not need extra lines to be fixed, as shown in
Listing~\ref{lst:fix}. Whereas, \emph{Broken Authentication} and vulnerabilities
such as the one presented in Section~\ref{sec:motivation} responsible by DoS
attacks are more difficult to fix (several code lines deleted and added).
Another example is the Apple authentication flaw discovered in $2017$
(CVE-$2017$-$13872$)\footnote{CVE-$2017$-$13872$ details available at
https://support.apple.com/en-us/HT208315 (Accessed on \today{})}, where any user
could log in as root with an empty password. Patrick Wardle examined the cause
of the issue\footnote{\emph{Why $<$blank$>$ Gets You Root} available at
https://objective-see.com/blog/blog\_0x24.html (Accessed on \today{})} and
conclude that the flaw was due to the introduction of high cyclomatic complexity
in a method used to verify the password. In sum, \emph{Cross-Site Scripting} and
\emph{Using Components with Known Vulnerabilities} should take part of the
developers' preoccupations along with the other patterns that explicitly hinder
maintainability because despite their low complexity these are patterns that
still produce a considerable negative impact on the maintainability. Real
efforts have been made in creating and incorporating coding standards and best
practices in software production. However, security is still far from perfection
and fixing vulnerabilities might hinder software maintainability. Thus, it is
important to provide tool support to developers to help them apply these
patterns without harming software maintainability. It is also important to redesign programming languages. \Rui{and redesign programming
languages?}

@@

All the security commits are described in English. This approach discards
security refactorings provided in other languages.\Rui{ehmm, isn't this
related to the first paragraph in the internal threat?}\Sof{I think we can close this one}

@@

Many studies have investigated the relationship between refactorings and
software quality. Previous work focused on object-oriented metrics \Rui{?}\Sof{Better?} has evaluated the
impact of refactorings and exhibited proof that quantifying effectively the
impact of refactoring on maintainability may help to choose the appropriate
refactoring type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of refactorings. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after refactorings. The differences
of maintainability between groups of refactored elements and non-refactored were
analyzed. In addition, the authors conclude that the source code elements
subjected to refactorings had significantly lower maintainability than the
elements not affected. Studying the evolution of maintainability issues during
the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes and faults. The present
work uses a model proposed by BCH and focuses solely on evaluating the impact of
patching vulnerabilities on software maintainability.

@@

This study also shows evidence that there are security patterns that need more attention than
others. \emph{Broken Autentication} issues, \emph{Denial-of-Service} attacks and
\emph{Memory Leaks} endure maintainability in $55.3\%$, $51.3\%$, and $47.5\%$,
respectively. \Sof{Check this par again}Following the implications of our work, we draw attention for the
need of code best practices documentation to
produce safe and maintainable code, proper authorization mechanisms design and
implementation, better tools to introduce in the CI/CD pipeline to help software
maintenance and new or improved programming languages to allow developers to
easily implement safety in their software without endangering maintainability.

As future work, this study may be extended in many directions: analyze which
guidelines are more likely to be affected by security-related refactorings;
investigate what are the most frequent vulnerabilities in each programming language\Rui{?}\Sof{Better?}; study if there is any correlation between software
maintainability and popularity of the project; expand our methodology with other
software quality properties; validate these findings with closed or private
software; and, expand this analysis to other quality standards.
